%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% acknowledge.tex: Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

I would like to say a huge thank you to my mom, Galina, who was the first person to introduce me
to the science world with small math questions and talks about stars and galaxies, to my 
dad, Sergey, who opened a gate to the world of physics competitions and provided the best resources 
to learn math and physics, and to my wife, Nina, who was always supporting and believing in me.
I could not achieve all of my goals without you. 

I want to say thank you to my adviser Gregory Pawloski for giving me this project, teaching and
sharing his wisdom in experimental physics research. Thank you to my former officemates
Susan, Dominick, Andrew and Jarrett with whom I shared the joy of working in NOvA and 
traveling to collaboration meetings, who helped me a lot when I struggled. Thank you
to our post-docs Matthew and Aaron for sharing their expertise and ideas.

A big thank you to my friends Yevhen, Mike, Sean, Yahor, Mitya, 
Olesya, Sergey, Timur, Vera for sharing with me countless hours of working through coursework,
papers, traveling around US and making my life easier and enjoyable. You are the best friends
one can dream of. 

And last but not least, I would like to thank you entire NOvA collaboration who make this work
possible, especially Alex Himmel, Bruno Zamorano and Jeremy Wolcott. You taught me skills
which were immensely important at the last stage of my work. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
